pdp: An R Package for Creating Partial Dependence Plots

### 2016-09-21 manuscript/greenwell.tex
reworded sentense
minor edit

The benefit of using \code{partial} is threefold: (1) it is a generic function that can be used for various types of fitted models (not just random forests), (2) it will allow for any number of predictors to be used (e.g., multivariate displays), and (3) it can utilize any of the parallel backends supported by the \CRANpkg{foreach} package \citep{foreach-pkg}; we discuss parallel execution in a later section. For example, the following code chunk assesses the joint effect of \code{lstat} and \code{rm} on \code{cmedv}. The results, which make use of various \code{plotPartial} options, are displayed in Figure~\ref{fig:pd_lstat_rm}.
\begin{example}
rwb <- colorRampPalette(c("red", "white", "blue"))
pd <- partial(boston.rf, pred.var = c("lstat", "rm"))
plotPartial(pd)
plotPartial(pd, contour = TRUE, col.regions = rwb)
plotPartial(pd, levelplot = FALSE, zlab = "cmedv", drape = TRUE,
            colorkey = TRUE, screen = list(z = -20, x = -60))
\end{example}
\strong{Note:} the default color map for level plots is the color blind-friendly matplotlib \citep{hunter-matplotlib-2007} 'viridis' color map provided by the \CRANpkg{viridis} package \citep{viridis-pkg}.

@@

Next, we use MARS to model ozone concentration as a nonlinear function of the eight meteorological variables plus day of the year; we allow for up to three-way interactions.
\begin{example}
ozone.mars <- earth(ozone ~ ., data = ozone, degree = 3)
summary(ozone.mars)
\end{example}
The model produced a generalized $R^2$ of $0.79$, similar to what was reported in \citet{breiman-1985-estimating}. A single three-way interaction was found involving the predictors
\begin{itemize}
  \item \code{wind}: wind speed (mph) at Los Angeles International Airport (LAX)
  \item \code{temp}: temperature ($^oF$) at Sandburg Air Force Base
  \item \code{dpg}: the pressure gradient (mm Hg) from LAX to Dagget, CA
\end{itemize}
To understand this interaction, we can use a PDP. However, since the partial dependence between three continuous variables can be computationally expensive, we will run \code{partial} in parallel.

>>>>

The benefit of using \code{partial} is threefold: (1) it is a generic function that can be used for various types of fitted models (not just random forests), (2) it will allow for any number of predictors to be used (e.g., multivariate displays), and (3) it can utilize any of the parallel backends supported by the \CRANpkg{foreach} package \citep{foreach-pkg}; we discuss parallel execution in a later section. For example, the following code chunk uses the random forest model to assess the joint effect of \code{lstat} and \code{rm} on \code{cmedv}. The results, which make use of various \code{plotPartial} options, are displayed in Figure~\ref{fig:pd_lstat_rm}.
\begin{example}
rwb <- colorRampPalette(c("red", "white", "blue"))
pd <- partial(boston.rf, pred.var = c("lstat", "rm"))
plotPartial(pd)
plotPartial(pd, contour = TRUE, col.regions = rwb)
plotPartial(pd, levelplot = FALSE, zlab = "cmedv", drape = TRUE,
            colorkey = TRUE, screen = list(z = -20, x = -60))
\end{example}
\strong{Note:} the default color map for level plots is the color blind-friendly matplotlib \citep{hunter-matplotlib-2007} 'viridis' color map provided by the \CRANpkg{viridis} package \citep{viridis-pkg}.

@@

Next, we use the multivariate adaptive regression splines (MARS) algorithm introduced in \citet{friedman-1991-mars} to model ozone concentration as a nonlinear function of the eight meteorological variables plus day of the year; we allow for up to three-way interactions.
\begin{example}
ozone.mars <- earth(ozone ~ ., data = ozone, degree = 3)
summary(ozone.mars)
\end{example}
The MARS model produced a generalized $R^2$ of $0.79$, similar to what was reported in \citet{breiman-1985-estimating}. A single three-way interaction was found involving the predictors
\begin{itemize}
  \item \code{wind}: wind speed (mph) at Los Angeles International Airport (LAX)
  \item \code{temp}: temperature ($^oF$) at Sandburg Air Force Base
  \item \code{dpg}: the pressure gradient (mm Hg) from LAX to Dagget, CA
\end{itemize}
To understand this interaction, we can use a PDP. However, since the partial dependence between three continuous variables can be computationally expensive, we will run \code{partial} in parallel.
